---
title: "Excercise_2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## Load the necessary libraries ################################################

library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times
library(ggpubr)       # To arrange multiple ggplots on the same page

#Task 1
## Import the downloaded csv ##################################################

ws_be <- read_csv("wildschwein_BE_2056.csv")
ws_be_sf <- st_as_sf(ws_be, coords = c("E", "N"), crs = 2056, remove = FALSE)

```

```{r}

# first overview
#producing data with timelag

ws_be_sf %>% 
  group_by(TierID) %>% 
  mutate(., timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC), units = "secs")) -> wildschwein_BE_sf
```

# wildschwein <- wildschwein_BE_sf %>% group_by(TierID) %>% mutate(timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC))) 



```{r}

# wildschwein$timelag  <- as.numeric(difftime(lead(wildschwein$DatetimeUTC), wildschwein$DatetimeUTC))
# wildschwein <- mutate(wildschwein,timelag = as.numeric(difftime(lead(DatetimeUTC),DatetimeUTC)))

  

#Task 2

# How many individuals were tracked?

wildschwein_BE_sf %>% count(TierID) 

count

# further exploration

wildschwein_BE_sf %>% 
  st_drop_geometry() %>% 
  group_by(TierID, TierName) %>% 
  summarise(pings = n(),
            start_time = min(DatetimeUTC), 
            end_time = max(DatetimeUTC),
            sample_interval = mean(timelag, na.rm = T),
            tracking_time = sum(timelag, na.rm = T)) %>% 
  mutate(tracking_time_days = days(day(seconds_to_period(tracking_time)))) 


```


# Were all individuals tracked concurrently or sequentially?
# For how long were the individual tracked? Are there gaps?


Creating visualisations
```{r}
ggplot(data = ws_be_sf) +
  geom_boxplot(aes(x = DatetimeUTC, y = TierID)) -> p1

p1

ggplot(data = wildschwein_BE_sf) +
  geom_histogram(aes(x = timelag, fill = TierID), binwidth = 100) +
  lims(x = c(0,15000))
  

#warum ist die Visualiserung so unpassend?


```


What is the temporal sampling interval between the locations?
```{r}
wildschwein_BE_sf %>%                         # Take wildschwein...
  group_by(TierID) %>%                        # ...group it by TierID
  summarise(                                  # Summarise the data...
    mean_timelag = mean(timelag,na.rm = T)    # ...by calculating the mean timelag
  )

# Answer:
#         002A = 1286
#         016A = 1412
#         018A = 1599
```



Task 3: Deriving movement parameters I: Speed
```{r}

wildschwein_BE_sf %>% 
  mutate(steplength = sqrt((E - lead(E,1))^2 + (N - lead(N,1))^2),
         speed = steplength/timelag)

# units are measured in m/s


```


Task 4: Cross-scale movement analysis
```{r}
#loading data

caro <- read_csv("caro60.csv")


caro_sf <- st_as_sf(caro, coords = c("E", "N"), crs = 2056, remove = FALSE)

?sf

caro_sf %>% 
  slice(seq(from=1, to = 200, by = 3)) -> caro_3
caro_sf %>% 
  slice(seq(from=1, to = 200, by = 6)) -> caro_6
caro_sf %>% 
  slice(seq(from=1, to = 200, by = 9)) -> caro_9

rm(caro3)

?remove


```



Calculate timelag, steplength and speed
```{r}

calculator <- function(df_input, interval){
  df_input %>% 
    mutate(timelag = as.numeric(difftime(lead(DatetimeUTC),DatetimeUTC), units = "secs"), 
           steplength = sqrt((E - lead(E,1))^2 + (N - lead(N,1))^2),
           speed = steplength/timelag,
           interval = interval) -> df_output
           
  
  return(df_output)
} 

calculator(caro_sf, "1 minute") -> caro
calculator(caro_3, "3 minutes") -> caro_3
calculator(caro_6, "6 minutes") -> caro_6
calculator(caro_9, "9 minutes") -> caro_9

rbind(caro, caro_3, caro_6, caro_9) -> caro_merge

```

```{r}

ggplot(data = caro_merge) +
  geom_line(aes(x = DatetimeUTC, y = speed, color = interval))

caro_merge %>% 
  filter(interval == "1 minute" | interval  == "3 minutes") %>% 
  ggplot(data = ., aes(x = E, y = N, color = interval, alpha = interval)) +
  geom_path() +
  geom_point() +
  scale_alpha_discrete(range = c(0.4, 1)) -> p1

p1

caro_merge %>% 
  filter(interval == "1 minute" | interval  == "6 minutes") %>% 
  ggplot(data = ., aes(x = E, y = N, color = interval, alpha = interval)) +
  geom_path() +
  geom_point() +
  scale_alpha_discrete(range = c(0.4, 1)) -> p2

p2

caro_merge %>% 
  filter(interval == "1 minute" | interval  == "9 minutes") %>% 
  ggplot(data = ., aes(x = E, y = N, color = interval, alpha = interval)) +
  geom_path() +
  geom_point() +
  scale_alpha_discrete(range = c(0.4, 1)) -> p3

p3




ggarrange(p1, p2, p3)


```


Task 5: Deriving movement parameters II: Rolling window functions
```{r}

library(zoo)

example <- rnorm(10)
rollmean(example,k = 3,fill = NA,align = "left")
##  [1] -0.5049544 -0.3781283 -0.1273157  0.4523153  0.4534811  0.8658536
##  [7]  0.5452571  1.0210648         NA         NA
rollmean(example,k = 4,fill = NA,align = "left")
##  [1] -0.25933381 -0.22379720  0.06456868  0.45949282  0.70918920  0.56899826
##  [7]  0.88605495          NA          NA          NA


library(zoo)
rollmean(caro$speed, k = 3)

```

